%%%%%%%%%%%%%%%%%%%%%%%%%

\label{section:example}
%%%%
Two examples are available in \fitter\ for benchmarking purposes.
The first example describes the default fit with HERA DIS inclusive data,
the second is a fit where all the available data in \fitter\ are fitted simultaneously.

\subsubsection{DIS inclusive only}
By default the \fitter\ steering files ({\tt steering.txt} and {\tt minuit.in.txt})
are set to fit the DIS inclusive cross section data. For this fit no any additional
configuration options are required.
The results of this example fit (total and partial $\chi2$, systematic shifts, 
data to theory comparison and histograms, see section 8.4 "Understanding the output" for more details) 
can be compared to the result provided for benchmarking in "examples".

%%%%
\subsubsection{All processes}
In order to run the second example with all data, the corresponding
{\tt steering.txt} from "input\_steering" has to be copied to the main directory:

\goodbreak                 
cp input-steering/steering.txt.ALLdata steering.txt
\\ \\
The user must make sure that all data sets as given in {\tt steering.txt.ALLdata} together with corresponding theory files have been downloaded  
%(\textcolor{red}{ ???}) 
before running this example. 
\\
Since the included sets contain various $ep$, $pp$, $p \overline p$ and fix target
data, it is necessary to have \fitter\ configured with applgrid, hathor and 
lhapdf options (corresponding shared library linking as explained in section~\ref{sec:install} is
required before configuration): 
\goodbreak
./configure --enable-applgrid --enable-lhapdf --enable-hathor
\\ \\
It is recommended to do "make clean" before each configuration. \\        
As in the previous case, the fit result can be compared to the one provided in "examples".



