
\fitter is an open source code and can be downloaded from \cite{herafitter:page}
together with the supporting documentation, README and additional fast grid theory files 
(see section \ref{sec:techniques}). 
The source code contains all relevant information to perform QCD fits with default 
HERA DIS data. The performance time depends on the fitting options and varies from 10 minutes 
(default settings) to several hours.
\fitter code is a combination of C++ and Fortran libraries with minimal 
dependencies, i.e. for the default fitting options no external dependences are required 
except \qcdnum evolution program \cite{qcdnum}. 
There are also cache options, fast evolution kernels, and usage of the OpenMP (Open Multi-Processing) 
interface which allows parallel applications of some of the heavy flavour scheme theory predictions in DIS. 
Information about the references and GNU public licence is provided 
together with the main source code as well. 

The \fitter project has successfully introduced into a wide variety
of tools to facilitate investigations of the HEP experimental data
and theoretical calculations. It provides a versatile interface
for understanding and interpreting new data and the derived PDFs.
%in the context of precision QCD theory. 
%\fitter\ has been successfully integrated in the high energy community 
%as a much needed means to provide understanding and interpretation 
%of new measurements in the context of QCD theory, a field limited by the precision of the PDFs.  
The \fitter platform not only allows the extraction of PDFs but also of theory 
parameters such as the strong coupling and heavy quark masses. The parameters and distributions are output with a
quantitative assessment of the fit quality with fully detailed information on 
experimental and theoretical uncertainties.
The results are also output to PDF \lhapdf grids that can be used to study predictions for 
SM or beyond SM processes, as well as for the study of the impact of 
future collider measurements (using pseudo-data).

So far the \fitter platform has been used to produce grids 
from the QCD analyses performed at 
HERA (\cite{h1zeus:2009wt}), and at the LHC, using 
measurements from ATLAS~\cite{atlas:strange,atlas:jets} (ATLAS PDF sets~\cite{atlas:grids}).
% to the QCD studies performed using pseudo-data to study the potential of possible future collider design such as LHeC.

For the following LHC analyses of SM processes 
the \fitter package was used:
%New results that have been based on the \fitter platform include 
%the following SM processes 
%studied at the LHC:  
inclusive Drell-Yan and $W$and $Z$ 
production~\cite{atlas:strange,atlas:hm,cms:strange};
inclusive jets~\cite{atlas:jets,cms:jets} production.
% and top measurements{\bf you need a reference for the top studies}
At HERA, the results of QCD analyses using \fitter are 
published for inclusive H1 measurements~\cite{h1:2012kk}
and the recent combination of charm production measurements 
in DIS~\cite{h1zeus:charm}.
The \fitter framework also provides a possibility to 
make impact studies for future colliders
as illustrated by the QCD studies that have been performed to 
explore the potential of the LHeC data~\cite{lhec:studies}.

A determination of the transverse momentum dependent gluon density using precision HERA
data obtained with \fitter has been reported in~\cite{Jung2014}. 

Recently a study based on a set of parton distribution functions 
determined with the \fitter program using HERA data was performed~\cite{hfcorrpaper}.
It addresses the issue of correlations between uncertainties for the LO,
NLO and NNLO sets. These sets are then propagated to study uncertainties 
for ratios of cross sections calculated at different orders in QCD and  
a reduction of overall theoretical uncertainty is observed.

