
\def\kt{\ensuremath{k_t}}
\newcommand{\Pmax}{p}
\newcommand{\CCFM}{CCFMa,CCFMb,Catani:1989sg,CCFMd}


%The proton PDFs are classically extracted from QCD fits by a measure of 
%the agreement between experimental data and corresponding theory models.
%During the fit procedure in the \fitter\ framework, the PDFs are
%parametrised at a starting scale $Q^2_0$ chosen to be below the charm 
%mass threshold and then evolved using coupled, integro-differential
%Dokshitzer-Gribov-Lipatov-Altarelli-Parisi 
%(DGLAP)~\cite{Gribov:1972ri,Gribov:1972rt,Lipatov:1974qm,
%Dokshitzer:1977sg,Altarelli:1977zs} evolution equations 
%as implemented in the QCDNUM~\cite{qcdnum} program in the $\overline{\text{MS}}$ scheme. 
%The evolution can be performed in the LO, NLO or NNLO accuracy~\cite{Curci:1980uw,Furmanski:1980cm}.
%
In this section the theoretical formalism for various processes available in \fitter\ is described.
%The PDFs are determined by measuring the agreement 
%the agreement between experimental data and corresponding theory models within 
%the DGLAP~\cite{Gribov:1972ri,Gribov:1972rt,Lipatov:1974qm,
%Dokshitzer:1977sg,Altarelli:1977zs} formalism.
%Models which are available in \fitter\ for various processes are described in the following text.



\subsection{Deep Inelastic Scattering Formalism and Schemes}
\label{dissection}


Deep Inelastic Scattering (DIS) data provide the backbone of any PDF fit.
The formalism that relates the DIS measurements to pQCD and the PDFs has been described
 in detail in many extensive reviews (see e.g.~\cite{disbook}) and will only be briefly recapped here.
DIS is a lepton scattering off the 
constituents of the proton by a virtual exchange of a NC 
or CC vector boson and, as a result, a scattered lepton and a 
multihadronic final state are produced.
The DIS kinematic variables are the negative squared four-momentum of 
the exchange boson, $Q^2$, the Bjorken $x$, 
%which can be related in the parton model to 
%the fraction of momentum carried by the struck quark, 
and the inelasticity $y$, where $y=Q^2/sx$ and $s$ is the squared centre-of-mass energy.
%$s = 4E_eE_p$ at HERA.
%which is the fraction of the energy 
%transferred to the hadronic vertex.
\\
%
The NC cross section can be expressed in terms of generalised structure functions:
\begin{eqnarray}
%  \nonumber
   \frac{d^2\sigma_{NC}^{e^{\pm} p}}{dxdQ^2}=\frac{2\pi\alpha^2}{xQ^4} 
     \big [ Y_{+} \tilde F_2^{\pm} \mp Y_{-}x \tilde F_3^{\pm} - y^2 \tilde F_L^{\pm} \big ],
 %\label{eq:NC not reduced}
\end{eqnarray}
where $Y_{\pm} = 1 \pm (1-y)^2$. The generalised structure functions $\tilde F_{2,3}$ 
can be written as linear combinations of the proton structure functions $F_2, F_{2,3}^{\gamma Z}$ 
and $F_{2,3}^Z$ associated to pure photon exchange terms, photon-$Z$ interference
terms and pure $Z$ exchange terms respectively. 
Structure function $\tilde F_2$ is the dominant contribution to the cross section, 
$x \tilde F_3$ is important at high $Q^2$ and $\tilde F_L$ is sizable 
only at high $y$. 
In the framework of pQCD the structure functions are directly related to the 
PDFs, i.e. in leading order (LO)  $F_2$ is the weighted momentum sum of quark and anti-quark distributions, 
$F_2 \approx x \sum e^2_q (q+ \overline q)$, $xF_3$ is related to their difference, 
$xF_3 \approx x \sum 2e_q a_q (q- \overline q)$ (where $a_q$ is the axial-vector 
quark coupling and $e_q$ the quark electric charge) and $F_L$ vanishes. 
At higher orders, terms related to the gluon density distribution
($\alpha_s g$) appear, in particular $F_L$ is strongly related to the low-$x$ 
gluon.
\\
The inclusive CC $ep$ cross section can be expressed 
in terms of another set of structure functions and in LO the $e^+p$ and $e^-p$ cross sections are sensitive to different quark flavour 
densities:
\begin{eqnarray}
%  \nonumber
    \begin{array}{rll}
   e^{+}:  & & \tilde \sigma_{CC}^{e^{+} p} = 
                x\overline U + (1-y)^2 xD  \\
   e^{-}:  & & \tilde \sigma_{CC}^{e^{-} p} = 
                xU + (1-y)^2 x\overline D.
    \end{array}
\end{eqnarray}
Here $U$ and $D$ denote the sum over up- and down-type quarks;
the latter include also strange and beauty quarks and 
the former charm quarks.
%
Beyond LO, 
%{\bf check this sentence if it makes sense}.
the QCD predictions for the DIS structure functions are obtained by convoluting 
the PDFs with the respective coefficient functions (hard process matrix elements).
%
The treatment of 
heavy charm and beauty quark production is a crucial point in these calculations
and several schemes exist:
\begin{itemize}
\item
In the Fixed Flavour Number (FFN) 
scheme~\cite{Laenen:1992, Laenen:1993, Riem:1995}
 only the gluon and the light quarks are considered
as partons within the proton and massive 
quarks (with mass $m_h$) are produced perturbatively in the final state.
The lowest order process is the fusion of a gluon in the proton
with a boson from the lepton to produce a heavy quark and an antiquark.
The modern series of PDFs that use this scheme as default are ABM and JR PDF groups.
\item
In the Zero-Mass Variable Flavour Number (ZM-VFN) scheme~\cite{ZMVFNpub} the
heavy quark densities are included in the proton for $Q^2$ values above a threshold $\sim m_h^2$
and
%for $Q^2>>m_h^2$ but 
are treated as massless in both the initial 
and final states. The lowest order process is the scattering
of a heavy quark in the proton with the lepton via (electroweak) boson exchange.
This scheme is expected to be reliable only in the region $Q^2 \gg m_h^2$.
This is the scheme that was used in the past by PDF groups.
\item
In the General-Mass Variable-Flavour Number (GM-VFN)
scheme~\cite{VFN} heavy quark production is treated for
$Q^2 \le m_h^2$ in the FFN scheme and for $Q^2 \gg m_h^2$
in a fully massive scheme. 
%in the ZM-VFN scheme with a suitable interpolation in between.
The modern series of PDF groups that use this scheme are MSTW, CT(CTEQ), NNPDF, and HERAPDF.
%This scheme is very popular and numerous variants exist.
\end{itemize}
All three schemes are available in \fitter\ . In the following
the implemented variants are briefly discussed.
% 
\begin{description}
\item \bf {FFN scheme:} \rm
In \fitter\ this scheme can be accessed via the 
\texttt{QCDNUM} implementation or through the interface to the open-source code \texttt{OPENQCDRAD} (as implemented by the ABM group)~\cite{openqcdrad:page}.
The latter implementation also includes the running mass definition of the heavy quark 
mass ~\cite{Alekhin:runm}.
The running mass scheme has the advantage of reducing the sensitivity of the DIS cross sections to
higher order corrections, and improving the theoretical precision of the mass definition. 
In \texttt{QCDNUM}, the calculation of the heavy quark contributions to DIS structure functions
are available at Next-to-Leading-Order (NLO) and only electromagnetic exchange contributions are taken into account.
In the ABM implementation the heavy quark contributions to CC structure functions are available 
and, for the NC case, the QCD corrections to the massive Wilson coefficients at Next-to-Next-to Leading Order (NNLO)
are provided at the best currently known approximation~\cite{SMoch:npb864}.
\item  
\bf {ZM-VFN scheme:} \rm 
The scheme is available for the DIS structure function calculation via interface to the \texttt{QCDNUM} package.
%
\item \bf {GM-VFN Thorne-Roberts scheme:} \rm
%\subsubsection{GM-VFN Thorne-Roberts scheme}
%
%The Thorne-Roberts (TR) scheme provides a smooth transition from the massive FFN
%scheme at low scales $Q^2<m_h^2$ to the massless ZM-VFNS scheme at  
%high scales $Q^2>>m_h^2$.
%There are two different variants of the TR schemes: TR standard 
%(as used in MSTW PDF 
%sets~\cite{Thorne:1997ga,Thorne:2006qt,MSTWpdf}) 
%and TR optimal~\cite{Thorne:6180}, with a smoother transition across the heavy quark threshold region.
%Both of these variants are accessible within the \fitter package at 
%NLO and NNLO.  
%
The Thorne-Roberts (TR) scheme~\cite{Thorne:1997ga} was designed to provide a smooth transition 
from the massive FFN scheme at low scales $Q^2 < m_h^2$ to the massless ZM-VFNS scheme at high scales $Q^2 \gg m_h^2$. 
However, the original version was technically difficult to implement beyond NLO, and was updated 
to the TR' scheme~\cite{Thorne:2006qt} which is simpler (and closer to the ACOT-scheme, see below).
There are two different variants of the TR' schemes: TR' standard (as used in MSTW PDF sets~\cite{Thorne:2006qt,MSTWpdf}) 
and TR' optimal~\cite{Thorne:6180}, with a smoother transition across the heavy quark threshold region. 
Both of these variants are accessible within the \fitter\  package at LO, NLO and NNLO.
%%%%
\vspace{0.1cm}
\item \bf {GM-VFN ACOT scheme:} \rm
The Aivazis-Collins-Olness-Tung scheme belongs to the group of VFN factorisation 
schemes that use the renormalization method of Collins-Wilczek-Zee (CWZ) \cite{CWZ}.
This scheme unifies the low scale $Q^2 < m_h^2$ and high scale $Q^2 > m_h^2$ regions; 
thus, it provides a smooth interpolation across the full energy regime. 
It is built upon the massive factorization theorem by Collins~\cite{CWZ} 
to incorporate the heavy quark masses for $Q^2 > m_h^2$; hence, it can be consistently applied 
order by order in the perturbation theory.
%
%This scheme involves a mixture of the $\overline{\text{MS}}$ scheme 
%for light and heavy (when the factorisation scale is larger than the heavy quark mass) partons
%and the zero-momentum subtraction renormalisation scheme for graphs with heavy quark lines 
%(if the factorisation scale is smaller than the mass of the heavy quark threshold). 
%% is this sentence below is important?
%%The DGLAP kernels and PDF evolution are pure $\overline{\text{MS}}$, 
%%therefore, the ACOT scheme is considered to be a minimal extension of the $\overline{\text{MS}}$ scheme.
Within the ACOT package, different variants of the ACOT scheme are available:
ACOT-Full, S-ACOT-$\chi$, ACOT-ZM, $\overline{\text{MS}}$ at LO and NLO. 
For the longitudinal structure function higher order calculations are also available. 
The ACOT-Full implementation takes into account the quark masses 
and it reduces to ZM $\overline{\text{MS}}$ scheme in the limit of masses going to zero, 
but it has the disadvantage that it is computationally intensive (addressed in 
section~\ref{sec:techniques}).

\vspace{0.1cm}
\end{description}

%%%%%%%%%%%
%\item \bf {Electroweak corrections for \texorpdfstring{$ep$}{ep} scattering:} \rm
Calculations of higher-order electroweak corrections to DIS scattering at 
HERA are available in \fitter, performed in the on-shell scheme 
where the gauge bosons masses $M_W$ and 
$M_Z$ are treated symmetrically as basic parameters together with the top, 
Higgs and fermion masses.
These electroweak corrections 
are based on the EPRC package~\cite{SpiesbergerPrivComm}.
The code provides the running of $\alpha$ using the most recent parametrisation
of the hadronic contribution to $\Delta_\alpha$ \cite{Jegerlehner}, as well as 
an older version from Burkhard \cite{Burkhard}.



\subsection{Drell Yan processes  in $pp$ or $p\bar p$ collisions}
\label{dysection}

%This section presents calculations of Drell Yan processes that can be used to 
%predict lepton pair production at the LHC or Tevatron.
The Drell Yan (DY) process
provides further valuable information about PDFs.
In $pp$ and $p\bar p$ scattering, the $Z/\gamma\*$ and $W$ production 
probe bi-linear combinations of quarks. 
Complementary information on the different quark densities
can be obtained from the $W$ asymmetry ($d$, $u$ and their ratio),
the ratio of the $W$ and $Z$ cross sections (sensitive to the flavor 
composition of the quark sea, in particular to the $s$ density), 
and associated $W$ and $Z$ production with
heavy quarks (sensitive to $s$ and $c$ quark densities).
%

Presently, the predictions for Drell-Yan and $W$ and $Z$ production are known 
to NNLO and $W$, $Z$ in association with heavy flavour quarks are known to NLO. There are several possibilities 
for obtaining the theoretical
predictions for DY production in \fitter. 
At LO an analytic calculation is available within the package and described
below: 

The LO DY triple differential cross section in
invariant mass \(M\), boson rapidity \(y\) and Centre of Mass 
lepton Scattering (CMS) angle \(\cos\theta\), for NC, 
can be written as~\cite{Drell:1970wh,Yamada:1981mw}:
\begin{align}
% \scriptstyle
 \textstyle
% \frac{\mathrm{d}^3\sigma}{\mathrm{d}M\mathrm{d}y\mathrm{d}\cos\theta} &= 
 \frac{d^3\sigma}{dM{d}y d\cos\theta} =  
 \frac{\pi\alpha^2}{3MS}\sum_{q}P_q \left[f_q(x_1,Q^2)f_{\bar{q}}(x_2,Q^2) 
 + (q\leftrightarrow\bar{q})\right],
\end{align}
where \(S\) is the squared CMS beam energy, \(x_{1,2} = \frac{M}{\sqrt{S}}\exp(\pm y)\), $f_q(x_1,Q^2)$ 
is the parton number density, and 
$P_q$ is a partonic cross section. 
%\begin{align}
%  P_q &=  e_l^2e_q^2(1+\cos^2\theta) \nonumber \\
%      &+  e_le_q\frac{2M^2(M^2-M_Z^2)}{\sin^2\theta_W\cos^2\theta_W
%          \big[(M^2-M_Z^2)^2+\Gamma_Z^2M_Z^2\big]} \nonumber \\
%      &    \big[aA_q(1+\cos^2\theta)+2bB_q\cos\theta\big] \nonumber \\
%      &+  \frac{M^4}{\sin^4\theta_W\cos^4\theta_W
%          \big[(M^2-M_Z^2)^2+\Gamma_Z^2M_Z^2\big]} \nonumber \\
%      &    \big[(a^2+b^2)(A_q^2+B_q^2)(1+\cos^2\theta)+8abA_qB_q\cos\theta\big].
%\end{align}
%Here \(\theta_W\) is the Weinberg angle, \(M_Z\) and \(\Gamma_Z\) are Z boson mass and 
%width, $a, b, A_q, B_q, e_l, e_q$ are electro-weak couplings.
%
%\begin{align}
% a & = -\frac{1}{4} + \sin^2\theta_W, \  b  = -\frac{1}{4}, \nonumber \\
% A_q & = \frac{1}{2}I_q^3-e_q\sin^2\theta_W, \ B_q  = \frac{1}{2}I_q^3, \ I_u^3  = -I_d^3 = \frac{1}{2},  \nonumber \\
% e_l & = -1, e_u = \frac{2}{3}, e_d = -\frac{1}{3}.
%\end{align}
\\
\\
The expression for CC  scattering has a simpler form:
\begin{align}
\frac{d^3\sigma}{dMdyd\cos\theta} &=
 \frac{\pi\alpha^2}{48S\sin^4\theta_W}
 \frac{M^3(1-\cos\theta)^2}{(M^2-M_W^2)+\Gamma_W^2M_W^2}  \nonumber \\
 & \sum_{q_1,q_2}V_{q_1q_2}^2f_{q_1}(x_1,Q^2)f_{q_2}(x_2,Q^2),
\end{align}
where \(V_{q_1q_2}\) is the CKM quark mixing matrix and \(M_W\) and \(\Gamma_W\)
are the \(W\) boson mass and decay width.

The simple form of these expressions allows the calculation of integrated
cross sections without the use of Monte-Carlo (MC) techniques which often 
introduce statistical fluctuations.
%This is particularly useful for PDF fitting purposes because
%statistical fluctuations are avoided in this case. 
In both NC and CC expressions PDFs
factorise as functions dependent only on boson rapidity \(y\) and
invariant mass \(M\), while
the integral in \(\cos\theta\) can be computed analytically.
%and integrations in \(y\) and \(M\) can be performed with the Simpson
%method. The \(\cos\theta\) parts are kept in the equation 
%explicitly because their integration is asymmetric for
%data in lepton \(\eta\) bins and also because of the need to apply 
%lepton \(p_{\perp}\) cuts.
%{\bf I think that this is an inappropriate level of detail about the LO DY calculation which could largely be replaced by references, however I have kept it for now, having moved it from the k-factor section where it did not belong}

The NLO and NNLO calculations are 
highly demanding
in terms of the computing power and time, and $k$-factor or fast grid techniques must be employed (see section~\ref{sec:techniques}
for details), interfaced to programs such as
\texttt{MCFM}~\cite{Campbell:1999ah,Campbell:2000je,Campbell:2010ff}, 
available for NLO calculations, or 
FEWZ~\cite{FEWZ} and DYNNLO~\cite{DYNNLO} for NLO and NNLO.
 

%The most abundant processes at the LHC are the production of the $W$ and $Z$ bosons, therefore measurements of the $W$ and $Z$ cross-sections are very precise. Their  LO decomposition in terms of quark distributions show strong 


%Alternatively, one can obtain the NLO predictions directly by using 
%APPLGRID or FASTNLO techniques, which rely on the factorisation theorem by 
%decoupling the hard scattering coefficients from PDFs.
%The hard scattering coefficients are calculated once and stored into a grid 
%for a given kinematic bin, speeding up the convolution process with the PDFs 
%and thus allowing to for fast QCD fits. 


%These methods are described in more detail in section \ref{sec:theory:jets}.
%An independent treatment for the electro-weak corrections is applied as the 
%independent k-factors, using packages such as SANC and FEWZ.

\subsection{Jet production in $ep$ and $pp$ or $p \bar p$ collisions}
\label{jetsection}
%In this subsection, the use of the factorisation formalism is fully exploited for the 
%calculations of the inclusive jets and dijet cross sections.
%This sections presents various fast calculational techniques for jet production based on
%the factorization formalism.

Jet production at high transverse momentum is sensitive to the high-$x$ gluon 
PDF (see e.g.~\cite{MSTWpdf}) and can thus increase the precision of the 
gluon PDF determination, which is particularly important for the Higgs production and searches for new physics.
Jet production cross sections are only currently known to NLO, although NNLO 
calculations are now quite advanced~\cite{nigel:2013,nigel:2010,Currie:2013dwa}. 
Within \fitter\ the programs such \texttt{MCFM} and 
\texttt{NLOJET++}~\cite{Nagy:1998bb,Nagy:2001fj} may be used for the 
calculation of jet production.
Similarly to the DY case, the calculation 
is very demanding in terms of computing power. 
%Although the concept is process independent, 
%The perturbative 
%coefficients are calculated by the \texttt{NLOJET++}, which is 
Therefore, to allow the possibility to include  $ep$, $pp$ or $p\bar p$ 
jet cross section 
measurements in QCD fits to extract PDFs and $\alpha_s$ fast 
grid techniques are used (see section~\ref{sec:techniques}).


%the perturbative
%coefficients have to be pre-computed in a PDF and $\alpha_s$ 
%independent way. For this reason, the fast grid tools to the theory calculations
%obtained with MCFM~\cite{Campbell:1999ah,Campbell:2000je,Campbell:2010ff} and NLOJET++~\cite{Nagy:1998bb,Nagy:2001fj}, 
%which are interfaced to the \fitter , are also exploited for the jet production . 



\subsection{Top-quark production in $pp$ and $p \bar p$ collisions}

Top-quark pairs ($t \bar t$) are produced at hadron colliders dominantly via $gg$ fusion 
and $q \bar q$ annihilation. Measured $t \bar t$ cross sections provide additional 
constraints in particular on the gluon density at medium to high values of $x$, 
on $\alpha_s$ and on the top-quark mass, $m_t$. 
Single top quarks are produced via electroweak interactions and single-top cross sections 
can be used, for example, to probe the ratio of the $u$ and $d$ densities in the proton 
as well as the b-quark PDF.
Precise predictions for the total $t \bar t$ cross section have become available 
to full NNLO recently~\cite{Czakon:2013goa}. They can be used within HERAFitter via an interface 
to the program \texttt{HATHOR}~\cite{Aliev:2010zk}. Differential $t \bar t$ cross sections and predictions 
for single-top production can be used with HERAFitter at NLO accuracy from 
\texttt{MCFM}~\cite{Campbell:2010ff,Campbell:2009ss,Campbell:2005bb,Campbell:2004ch,Campbell:2012uf} 
in combination with fast grid techniques.

%\subsection{Cross Sections for \texorpdfstring{$t\bar{t}$}{t-tbar} production in $pp$ or $p\bar p$ collisions}
%
%This provides the possibility to use top production to
%constrain the gluon density in the proton. Calculations are available to NLO precision with  
%MCFM and to approximate NNLO precision with the HATHOR program~\cite{Aliev:2010zk}. These are both available within \fitter\. 
%Version 1.3 of HATHOR includes the exact NNLO for $q \bar q \to t \bar t$ \cite{Baernreuther:2012ws}
%as well as a new high-energy constraint on the 
%approximate NNLO calculation obtained from
%soft-gluon resummation \cite{Moch:2012mk}.
%The use of these programs also requires fast grid techniques.
