
\label{sec:pdfparam}
%%%%%%%%%%%
%Different parametrisation for PDFs at the starting scale present in \fitter\ are described in this section.
%It starts from various standard functional forms (\ref{sec:stdform}), it follows with the bi-log-normal functional form (\ref{sec:log}), and it extends to more exotic forms based on generalised polynomial such as Chebyshev (\ref{sec:cheb}).

In this Section the different possible parametrisations for the PDFS at the staring scale which are implemented in the \fitter\
are described. Sec.~\ref{sec:stdform} covers standard functional forms, Sec.~\ref{sec:log}  covers the bi-log normal functional 
for and Sec.~\ref{sec:cheb} covers more exotic forms based on generalised polynomials such as the Chebyshev polynomials.

\subsubsection{Standard Functional form}
\label{sec:stdform}
%%%%
The term standard functional form is understood to refer to a simple polynomial 
that interpolates between the low and high $x$ regions:
\begin{equation}
 xf(x) = A x^{B} (1-x)^{C} P_i(x),
\label{eqn:pdf_std}
\end{equation}
Standard forms are commonly used by PDF groups.

The notation used throughout this text reflects the 
notation used in the code.

%%%%
\begin{description}

%%%%
\item \bf{HERAPDF style}\rm

%\subsubsection
%%%%
 The parametrised PDFs at HERA are the valence distributions
 $xu_v$ and  $xd_v$,  the gluon distribution $xg$, and the $u$-type and $d$-type sea 
$x\bar{U}$, $x\bar{D}$, where $x\bar{U} = x\bar{u}$, 
$x\bar{D} = x\bar{d} +x\bar{s}$. 
The following standard functional form is used to parametrise them
\begin{equation}
 xf(x) = A x^{B} (1-x)^{C} (1 + D x + E x^2),
\label{eqn:pdf}
\end{equation}
%
where the normalisation parameters, $A_{uv}, A_{dv}, A_g$,  
are constrained by the number sum-rules and the momentum sum-rule.
The $B$ parameters  $B_{\bar{U}}$ and $B_{\bar{D}}$ are set equal,
 $B_{\bar{U}}=B_{\bar{D}}$, such that 
there is a single $B$ parameter for the sea distributions. 
%
The strange quark distribution 
is already present at the starting scale and 
%
it is assumed that
$x\bar{s}= f_s  x\bar{D}$ at $Q^2_0$. 
The  strange fraction is chosen to be $f_s=0.31,$ which is
consistent with determinations 
of this fraction using neutrino induced di-muon production. 
%
In addition, to ensure that $x\bar{u} \to x\bar{d}$ 
as $x \to 0$,  
$A_{\bar{U}}=A_{\bar{D}} (1-f_s)$.
%
The $D$ and $E$ parameters are introduced one by one until no further improvement in $\chi^2$ is found.
This procedure allows for more flexibility when adding more precision data into the fit, for example when adding HERA-II data. 
The best fit  results in a total of $10$ free parameters when performing fits to solely HERA I data 
(fits are then referred to as HERAPDF1.0), and of 13 free parameters when adding preliminary HERA II data on top 
(fits are then referred to as HERAPDF1.5).

In HERAPDF fits the assumption is made that the strange sea quark density follows the same shape as the down quark density. 
However, and alternative functional form is provided to parametrise the strange density  at the starting scale:
\begin{equation}
xs(x)=f_s\frac{1}{1+\tanh(-(x-x_{hs}h_{hr}))},
\end{equation}

with $x_{hs} = 0.07$ and $h_{hr} = 20$, corresponding to a sharp turn on of the strange density at $x\sim 0.07$.
This form is inspired by the HERMES analysis \cite{hermes} which measure semi-inclusive production of the strange mesons.

\item \bf{Strange style}\rm


For studies of the strange quark sea density, the parametrisation of $x\bar{D}$ is replaced by a sum of $x\bar{d}(x)+x\bar{s}(x)$ parametrised densities:
\begin{equation}
x\bar{s}(x)=\frac{fs}{1-fs} A_{\bar{d}}x^{B_{\bar{s}}}(1-x)^{C_{\bar{s}}} .
\end{equation}

\item \bf{Flexible style}\rm
%%%%

Flexible style is the extension of the ``HERAPDF style'' by allowing $2$ extra free parameters for every PDF distribution, 
namely the $D$ and $E$ parameters which give flexibility in the medium $x$ range. 
This can be used to study the sensitivity of the data to the PDF description. In this case the total number of free parameters is $22$.

%Strange quark PDF is the least determined PDF from the l

\item \bf{CTEQ style}\rm

%%%%

\begin{equation}
 xf(x) = a_0 x^{(a_1+n)} (1-x)^{a_2} e^{a_3x} (1 + e^{a_4} x + e^{a_5} x^2),
\label{eqn:pdf_cteq}
\end{equation}
%
\end{description}

\subsubsection{Bi-Log-Normal Functional Form}
\label{sec:log}
A bi-log-normal distribution is proposed by \cite{AndreSchoening} to parametrise the $x$ dependence of the PDFs.
%parton distribution function of the proton.
This parametrisation is motivated from multi-particle statistics. The follwing parametrisation is proposed as a general ansatz.
\begin{equation}
xf(x)=x^{p-b\log(x)}(1-x)^{q-\log(1-x)}.
\label{eq:AS}
\end{equation}
This function can be regarded as a generalisation of parametrisations commonly used by global fit groups.
In order to satisfy the QCD sum rules this parametric form requires numerical integration.

\subsubsection{Chebyshev Polynomial Functional Form}
\label{sec:cheb}
A flexible Chebyshev polynomial based parametrisation can be used for the gluon and sea densities. The polynomials
use $\log x$ as an argument to emphasize the low $x$ behavior. 
The parametrisation is valid for $x>x_{min} = 1.7\times 10^{-5}$. The PDFs are multiplied
by $1-x$ to ensure that they vanish as $x\to 1$. The resulting parametric form is 
\begin{eqnarray}
x g(x) &=& A_g \left(1-x\right) \sum_{i=0}^{N_g-1} A_{g_i} T_i \left(-\frac{\textstyle 2\log x - \log x_{min} } {\textstyle \log x_{min} } \right)\,, \label{eq:glu} \\
x S(x) &=& \left(1-x\right) \sum_{i=0}^{N_S-1} A_{S_i} T_i \left(-\frac{\textstyle 2\log x - \log x_{min} } {\textstyle \log x_{min} } \right)\,. \label{eq:sea} 
\end{eqnarray}
Here the sum over $i$ runs up to $N_{g,S}=15$ order Chebyshev polynomials of the first type $T_i$ for
the gluon, $g$, and sea-quark, $S$, density, respectively. 
The normalisation $A_g$ is given by the momentum sum rule.

The advantages of this parametrisation are that the momentum sum rule can be evaluated analytically 
and that for $N \ge 5$ the fit qulaity is already similar
to a standard Regge-inspired parametrisation with a similar number of parameters.

%==========================================
\subsubsection {Diffractive parametrisation Functional Form}

\begin{description}
\item \bf{Pomeron parametrisation}\rm

\newcommand\AP {{\cal P}}

The Pomeron is parametrised at the initial
$Q_0^2$ in terms of two singlet distributions,
$f_{g}$ and $f_{+}$, which evolve as follows:
\begin{subequations}

\label{singlet}
\begin{eqnarray}
\frac{d}{dt}f_{+} &=&
\asotp\left[\AP_{\rm FF} f_{+} +\AP_{\rm FG}f_g
\right]
\\
\frac{d}{dt}f_{g} &=&
\asotp\left[
\AP_{\rm GF} f_{+} +\AP_{\rm GG}f_g
\right]
\end{eqnarray}
\end{subequations}

As $\Pom$ is neutral, $f_{q} = f_{\bar q}$ for each flavour $q$.
Assuming that all light quark PDFs are equal
\begin{equation}
f_d = f_u = f_s
\,,
\end{equation}
we have
\begin{subequations}
\label{eq:pm}
\begin{eqnarray}
f_{q-} &\equiv& 0
\\
f_{q+} &\equiv& 2 f_q
\end{eqnarray}
\end{subequations}

At $n_f = 3$
\begin{equation}
\label{eq:fq3}
f_{q+} = f_{+}/3,\; q = d,u,s
\,.
\end{equation}
% \ifFullVer
i.e.
% \begin{equation}
% \Sgl q = 0,\; q = d,u,s
% \,,
% \end{equation}
% where
\begin{equation}
\Sgl q \equiv f_{q+} - \frac{1}{n_f}f_{+}
 = 0,\;\mbox{for}\; q = d,u,s
\,.
\end{equation}
% \fi

This gives all PDFs for the FFNS, while for VFNS 
$f_{h+}$ for $h=c,b,t$ are generated dynamically above the respective
transition scales $Q_h^2$.
Hence at $n_f > 3$ the singlet has contributions from the heavy quarks
and we get non-trivial nonsinglet distributions $\Sgl{h}$ satisfying
\begin{equation}
\label{eq:nsevol}
\frac{d}{dt}\Sgl h = \asotp\, \AP_{(+)} \Sgl h
\end{equation}

\end{description}
%\subsection {Parametrization at \texorpdfstring{$Q_0^2$}{Q0}}
%\subsubsection {Parametrization at {$Q_0^2$}}
{\bf Parametrisation at {$Q_0^2$}} \\
%-----------------------------------------------------------
\label{sec:Par}

Full PDFs are given in analogy to Eq.~\ref{eq:FD3}
\begin{equation}
f_k^{D(3)}(\beta,Q^2,\xi) =
\hat\Phi_\Pom(\xi)\, f^\Pom_k(\beta,Q^2)
+
\Phi_\Reg(\xi)\, f^\Reg_k(\beta,Q^2)
\end{equation}
where $\hat\Phi_\Pom \equiv \Phi_\Pom/A_\Pom$,
with the fluxes given by Eq.~\ref{eq:intFlux} and Eq.~\ref{eq:flux}.

The Pomeron PDFs are parametrised as
\def\Cini#1#2{A^{(#1)}_#2}
\begin{equation}
\label{eq:fP0}
f^\Pom_N = \Cini N1  x^{\Cini N2} (1-x)^{\Cini N3}
%   \left(1 + \Cini N4 x \right)
  \; \exp\left(-\frac{d}{1.00001-x}\right)
\,,
\end{equation}
where the `dumping factor' $d$ is taken as 0.01 or 0.001.
$N = \mathrm G$ for gluon and $N = \mathrm S$ for `singlet': $f_{\rm S} \equiv f_+(n_f=3)$,
cf. Eq.~\ref{eq:fq3}.



% The Reggeon PDFs $f^\Reg_k$ are taken from pion.

%\subsubsection {HERAFitter parameters}






